---
title: "R Notebook"
output: html_notebook
Date: 1/8/2021
Author: "SREANG VUTHY"
---

### 1. Write a R code to Generate 6X6 matrix (A) using the random numbers between 1 to 100 and generate another 6x6 matrix (B) using random numbers between 500 to 600.
a. Find the range of matrices A and B.
b. Find the max value from each row of the matrices A and B.
c. Find the diagonal values of matrices A and B.
d. Multiply matrices A and B
e. Replace the 4th row of Matrix A with diagonal values of matrix B
f. Multiply matrices A and B and store the values on X.
g. Divide the matrix A with matrix B.

```{r}
# 1. create matrix A and B
a <- sample(1:100)
b <- sample(500:600)
A <- matrix(a, nrow = 6, ncol = 6)
B <- matrix(b, nrow = 6, ncol = 6)
A
B

# a. Find the range of matrices A and B
range_a <- range(A)
range_b <- range(B)
range_a
range_b

# b. Find the max value from each row of the matrices A and B
install.packages("matrixStats")
library(matrixStats)
max_a <- rowMaxs(A, na.rm = TRUE)
max_b <- rowMaxs(B, na.rm = TRUE)
max_a
max_b

# c. Find the diagonal values of matrices A and B
diag_a <- diag(A)
diag_b <- diag(B)
diag_a
diag_b

# d. Multiply matrices A and B
multiply_ab <- A %*% B
multiply_ab

# e. Replace the 4th row of Matrix A with diagonal values of matrix B
rep_a <- rep(A, diag_b)
rep_a

# f. Multiply matrices A and B and store the values on X
X <- A %*% B
X

# g. Divide the matrix A with matrix B
divid_ab <- A/B
divid_ab
```


### 2. Create a table using dataframe/tibble with the following categories(a to f) (10 marks)

a. First name,
b. Second Name,
c. Age,
d. Occupation,
e. Place,
f. Random numbers( between 1 to 10)

And input at least 10 samples(observation) and do the following

g. Extract the last column (f)
h. Find the sum, mean and length of the extracted column

```{r}
# a to f: create dataframe
a <- c("Vuthy", "Samnang", "Ratanak", "Channy", "David", "Zevo", "Sokmeng", "Moonlight", "Kimteng", "Dara")
b <- c("Sreang", "Set", "Seng", "Pong", "Kat", "Thy", "Pech", "Kim", "Chy", "Kruy")
c <- c(21, 21, 20, 23, 21, 34, 23, 22, 20, 24)
d <- c("Student", "Student", "Student", "Student", "Student", "Progammer", "Teacher", "Engineer", "Student", "Student")
e <- c("Speu", "Takeo", "Phnom Penh", "Kompong Cham", "Svay Rieng", "Takeo", "Batombong", "Kompong Cham", "Svay Rieng","Takeo")
f <- c(1,2,3,4,5,6,7,8,9,10)
my_data_frame <- data.frame(First_Name = a, Second_Name = b, Age = c, Occupation = d, Place = e, Random_Number = f)
my_data_frame

# g. extract the last column (f)
extracted_column <- my_data_frame$Random_Number
extracted_column

# Find the sum, mean and length of the extracted column
sum_ex <- sum(extracted_column)
sum_ex

mean_ex <- mean(extracted_column)
mean_ex

length_ex <- length(extracted_column)
length_ex

```

### 3. Create a .txt file (Minimum : 150 words) with the latest news (In English) in cambodia from the official news agency’s website(Note: Include the reference link) and do the following

a. Read the .txt file
b. Perform the Tokenization and count the words
c. Identify the most repeated word and replace that word with “KIT”
d. Perform the Sentiment analysis
e. Plot the graph between “words count” and “words” using geom_col()

e.1. X-axis title =“Number of words”, Size=10, Bold, Text in Blue color,
Center, space between graph and title = 20.

e.2. Y-axis title =“List of words”, Size=10, Bold, Text in Red color,
Center, space between graph and title = 20.
e.3. Graph title =”Text Mining using R”,Size = 25, Bold, Center, Text in
Yellow color, space between graph and title = 25.

```{r}
library(tokenizers)
# a
my_data <- read.delim("News.txt")

# b
words_data <- paste0(my_data)
tokens <- tokenize_words(words_data)
count <- count_words(words_data)
count

# c
most_words <- names(table(tokens))[as.vector(table(tokens))==max(table(tokens))]
most_words_count <- max(table(tokens))
subs <- gsub("oil", "KIT", tokens)
most_words
most_words_count
subs

# d
library("syuzhet")
library("ggplot2")
sentiment_analysis <- get_sentiment(tokens, method="syuzhet")
summary(sentiment_analysis)
# e


```



### 4. Find the assigned Dataset from the list below and perform the following (20 Marks)
a. Import and Clean the dataset.
b. Find the x and y variables(axis) from the assigned dataset and give the reason for
selecting the x and y axis.
c. Plot the graph using geom_col() and perform the linear model in statistics.
d. Perform the facet with the available columns(Choose the any two variables from the
columns)
e. Use the theme layer to do the following (Note: Based on dataset, name the x and y axis &amp;
graph)
e.1. x-axis title =“________”, Size=10, Bold, Text in Green color, Center, space
between graph and title = 20.
e.2. y-axis title =“________”, Size=10, Bold, Text in Yellow color, Center, space
between graph and title = 20.
e.3. Graph title =”_________”,Size = 25, Bold, Center, Text in Blue color, space
between graph and title = 25.

Dataset Links: "https://github.com/fivethirtyeight/data/blob/master/media-mentions-
2020/cable_weekly.csv"

*NOTE: I have download that dataset in my computer then I import it through below path(In my computer)

```{r}
# a. Import and Clean the dataset.
library("tidyverse")
dataset <- read_csv("/Users/VUTHY/Downloads/cable_weekly.csv")
dataset

#is.na(dataset)
# the dataset is already clean there is no NA so no need to clean again.

# b. Find the x and y variables(axis) from the assigned dataset and give the reason for
# selecting name as X and total_clips as Y because it would be easy to view name in X and total_clips in Y to easy to understand which name with the total_clips in graph.
plot <- ggplot(dataset, aes(x=name, y=total_clips)) + 
    xlab("Name") +
    ylab("Total_Clips")

# c. Plot the graph using geom_col() and perform the linear model in statistics.
# using geom_col
plot <- plot+geom_col()

#perform linear model in smooting
plot <- plot+geom_smooth(method="lm")
plot

# d. Perform the facet with the available columns(Choose the any two variables from the columns)
# use facet with 2 variable by the total_clips and date
plot+facet_grid(total_clips ~ date)

# use facet with 1 variable by the matched_clips
plot <- plot+facet_wrap(~total_clips)
plot
# because it diffcult to view 2 variable so choose only 1 variable

# e. Use the theme layer to do the following (Note: Based on dataset, name the x and y axis graph)
#e.1. x-axis title =“________”, Size=10, Bold, Text in Green color, Center, space between graph and title = 20.
#e.2. y-axis title =“________”, Size=10, Bold, Text in Yellow color, Center, space between graph and title = 20.
#e.3. Graph title =”_________”,Size = 25, Bold, Center, Text in Blue color, space between graph and title = 25.
plot <- plot +
  ggtitle("Cable Weekly in 2020") + theme(
  axis.title.x=element_text(size=10, color="green", face="bold", margin=margin(t=20,r=20,b=0,l=0)),
  axis.title.y=element_text(size=10, color="yellow", face="bold", margin=margin(t=0,r=20,b=0,l=20)),
  plot.title=element_text(size=25, color="blue", face="bold", margin=margin(t=25,r=0,b=25,l=0))
  )
plot

ggsave(filename="cable_weeklyGraph.png", units="cm", width=25, height=25)

```

