---
title: "R Notebook - Unit1"
output: html_notebook
---

## R Language
R was created by Ross Ihaka and Robert Gentleman on the Auckland university of New Zealand.
R is a programming language, and developing environment used for statistic computing, graphic and processing/ analyzing data during data mining process (clean, analyze and graph your data).


## Basic R
```{r}
# set working directory
setwd(dir = "C:/Users/VUTHY/Documents/R")

# get working directory
getwd()
```
# Data types and variables 
- objects are variables while class is their type
- auto defined class by the value assigned to the object
- value assignment : <- or =
- objects (variables) : '_', '.', starts with alphabets, special char not allowed and assigned values at the time of declaration

### Basic Atomic object class
* Character (ex: 'Data')
* Numeric (ex: 1,3,5.67,34)
* Integer (ex: 10L,4L,4L)
* Complex (ex: 3+1i)
* Logical - True/False (ex: True, False, NA)

```{r}
# get user input
age = as.integer(readline(prompt = "Input your age: "))

# output
print(age)
```
### Data Structure
* Homogeneous : all elements are of the same data type(class) -> Vector, Matrix, Array
* Heterogeneous: Elements can be of more than one data type(class) List, DataFrame

```{r}
# Vector : indexed starting at 1, 
vec_1 = c(1: 5)
vec_1
vec_2 = c(1, 2, 3, 4, 5)
vec_2

# Sequence
seq = seq(from=1, to=10, by=0.5)
seq
# Repeat
rep = rep("xxxx",3)
rep
```

```{r}
# Subset
x = c(1:5)
x
x
sub_x = x[1:3]
sub_x

# Inner Product
inner_x = x %*% x
inner_x
# inner_x

# Outer Product
outer_x = x %o% x
outer_x
```
# List 
list is a set of objects which might belong to another class 
```{r}
x<-list("KIT", NA, 1, 3.5)
x[3]
test_list = list( a = c(1, 2, 3, 4), b = TRUE, c ="Hello")
test_list
test_list$a
test_list[3]
```

# matrix is a collection of multiple vectors
```{r}
# Matrix
matrix1<- 1:15
matrix1

# set dimension of matrix
dim(matrix1) = c(3, 5)
matrix1
dim(matrix1)

# using matrix function
row_col = 1:9
matrix2 = matrix(row_col, 3, 3, byrow = FALSE)
matrix2

# using rbind
row1 = 1:10
row2 = 10:21
matrix3 = rbind(row1, row2)
matrix3
# using cbind
col1 = 1:10
col2 = 11:20
matrix4 = cbind(col1, col2)
matrix4
```

```{r}
# Matrix Calculation
a<- 1:9
b<- 2:10
x<-matrix(a,3,3,byrow=TRUE)
y<-matrix(b,3,3,byrow=TRUE)
x
y
# Matrix multiplication
x %*% y

x*x

# Matrix Inverse
#solve(y)
library(MASS)
print(ginv(y))

print("-----")
# Transpose
t(x)

print("-----")
# Diagonal
diag(x)

print("-----")
diag(3,3,3)

print("-----")
diag(1:4)

print("-----")
eigen(x)

```
# Data Frames are used to store data in a table format
 - each column can only contain objects of the same class
```{r}
# DataFrame
m<-c("Kimseng", "Ratana", "Phanith")
l<-c(23,24,25)
q<-c(158,160,171)
exm<-data.frame(name=m, age=l, height=q)
exm
```
# Data import

- read.table, read.csv
- reads SPSS, Stata and SAS files(tidyverse)
- readxl reads excel files(both .xls and .xlsx)
```{r}
install.packages("tidyverse")
library(tidyverse)
a = read_csv("C:/Users/VUTHY/Downloads/us_counties_covid19_daily.csv")
a
```
# Data Export 
```{r}
install.packages("xlsx")
library(rJava)
library(xlsx)
# library(readr)
write_csv(exm, "C:/Users/VUTHY/Downloads/covidtest.csv", append=FALSE)
write.xlsx(exm, "C:/Users/VUTHY/Downloads/covidtest1.xlsx", append=FALSE, sheetName="C:/Users/VUTHY/Downloads/covidtest1.xlsx")
```

# Control Structure 
- conditional statements : if/else, elseif, ifelse()
  if(){
  
  }else{
  
  }
  ifelse(---, 1(yes), 0(no))
- Loops: for, repeat(endless loop), while
  for (object in vector) {
  
  }
  repeat{
    if(){
      break
    }
  }
  while(){
  
  }
- Functions are defined, using keyword function
  name<- function(a, b, ...){
  
  }
```{r}
# if/else
num = 6
if(num == 6){
  print("YES")
}else{
  print("NO")
}

#ifelse
ifelse(num == 6, "YES", "NO")
```
  
  

# Data Manipulation and Visualization

* Data manipulation is the changing of data to make it easier to understand or be more organized, to display information to users in a more meaningful way; (Data Wrangling or cleaning)
* Learn how to explore, filter, reorganize and process your data;
* Most of the datasets are not  clean and tidy so we need to prepare the datasets before performing the statistical analysis
* Rules for Data Tidy:
  * Variable - Own column
  * Observation - Own Row
  * Variable - Own cells
* ***dplyr*** is a package that makes data manipulation easy; It consists of five main verbs: 
  * filter() 
  * arrange()
  * select()
  * mutate()
  * summarise()
  * group_by()

### Subset
* Check the observations(Row) and variables(column)
  * ncol(tablename)
  * nrow(tablename)
  * dim(tablename)
* To keep first few and last observations
  * head(tablename, n=5)
  * tail(tablename,n=5)
* Random sample of observations
  * sample_n(tablename,n, replace=FALSE)
* select observations & variables
  * table_name[c(3:5), ]  (or) table_name[ , c(3:5)] 
* Remove observation
  * table_name[-c(2:4), ]  (or) table_name[ , -c(2:4)]
* based on variables
  * table_name$variables
```{r}
x <- ggplot2::diamonds
x
names(x)
help(x)
ncol(x)
nrow(x)
dim(x)
head(x, n=5)
tail(x, n=5)
#sample_n(x,5, replace=FALSE)
x[c(3:5,15,18), 4]
x[3,4]
x[c(3:5),]
x[,c(3:5)]
#remove
x[-c(3:5),]
x[,-c(3:5)]
x$cut
```
# Missing Values 
- Missing values (NA - "Not Applicable") are often problematic for many analyses
- mean of a series or variable with at least one NA will give a NA

```{r}
is.na(x)  #check missing value
which(is.na(x)) #return the position of missing values(by cell)
apply(is.na(x), 2, which)  #get the positions of each column in your data set
na.omit(x)  #delete the observation
h<-mean(x$carat, na.rm = TRUE)
g<-impute(x)  #instead of removing, replace by some values such as the median of the variable
```
- filter is a function that filter out rows that meet certains conditions
- arrage : sorting data frames 
- select : select the variables 
- mutate : manipulating and creating variable (column)
- transmute : will return only the column that u have manipulated
  . rank(), min_rank(), dense_rank(), percent_rank() to rank
  . log(), long10(), to take log of a variable
  . cumsum(), cummean()
  . row_number() if u need to create row numbers
  . lead(), lag()
- summarise() or summarize() : summarizing data (mutation)
- group_by() : entire dataset to operating on it group-by-group
- rename() : rename the variable

```{r}
m<-filter(x, price > 10000)
m
q<-arrange(x, desc(cut))
q
f<-dplyr::select(x, price, color)
f
h<-dplyr::select(x, cut, starts_with("a"), ends_with("b"), contains("h"), matches("j"))
h
k<-mutate(x, kkkk=cut)
k
t<-transmute(x, kkkk=price)
t
s<-summarise(x, f=log(price), g=mean(color))
s
g<-group_by(x, price)
g
uuu<-rename(x, xxxx=cut)
uuu
```

```{r}
x %>% filter(price < 400)  %>%  group_by(price) %>% summarise( mea = mean(price, na.rm = T))
```

# Visualization
1. Data : our datasets
2. Aesthetics : Axes, plot, positions, encodings?
  - variable mapping like aes(x=var1, y=var2, color="red")
3. Scale : scale values, represent multiple values?
  - scale_x_continuous(limits=c(0,5)) +
  scale_y_continuous(limits=c(0.20))
4. Geometric objects : line, bar, points?
  - type of plot 
    0 dimensions - point, text
    1 dimensions - path, line
    2 dimensions - polygon, interval
  Ex: ggplot(...) + geom_point()
  for bar_chart : geom_bar(stat="identity", position="dodge") (to avoid overlap)
5. statistics : mean, quantitile, confidence intervals?
  - transform the data, generally by summarizing the info
  - a stat is a func that takes in a dataset as the input and return output 
  - stat can add new var to the original dataset, or create an entirely new dataset
  Ex: geom_smooth(method="lm")
6. facets : create subplots based on multiple dimensions
  - can be used to split the data up into subsets of the entre dataset
  - describes which variables should be used to split up the data and how they should be arranged
  ex: ggtitle("title"), xlab("title"), ylab("title")
  facet_wrap(~variable) - subplot into column
  facet_grid() - subplots to row and column 
7. coordinate system : cartesian, polar?
  - map the position of objects onto the plane of the plot and controls how the axes and grid lines are drawn 
  Ex: coord_catesian()

***save: ggsave(filename="nyrattsart.png", units="cm", width=25, height=25)

```{r}
#tidy dataset 
data<-ggplot2::diamonds
data
```
```{r}
#optional: scale, default grph: cartesian_coor
#graph aes, geom
ggplot(data, aes(x=carat, y=price, color=clarity)) +
#geometry
    geom_point(size=3, alpha = 1/3) +
#scale
    scale_x_continuous(breaks=seq(0.5, 0.5)) +
    scale_y_continuous(breaks=seq(0,18000,1000)) +
#coordinate system
    coord_polar() +
#give title
    xlab("weight of the diamonds") +
    ylab("price of diamonds") +
    ggtitle("price of over 50,000 round cut diamonds") +
#statitic
    geom_smooth(method="lm") +
#facet - subplotting (make to small small plot instead of previous plot): specific value and need to understand the column
    facet_grid(cut ~ color) +
#extra
#theme layer : change the theme of graph (want to change to black and white for more clean or sth)
    theme_bw() +
#increate the size and make to bold ...
    theme(axis.title.x=element_text(size=15, color="blue", face="bold", margin=margin(t=20,r=20,b=0,l=0)),
          axis.title.y=element_text(size=15, color="red", face="bold", margin=margin(t=0,r=20,b=0,l=20)),
          plot.title=element_text(size=15, color="green", face="bold", margin=margin(t=20,r=0,b=20,l=0)),
#box color
          strip.background=element_rect(color="yellow", fill="pink"))
```

